---
title: "Advanced Analysis: Imaginary Network Motifs"
author: "Sima Najafzadehkhoei"
date: today
format: html
vignette: >
  %\VignetteIndexEntry{Advanced Analysis}
  %\VignetteEngine{quarto::format}
  %\VignetteEncoding{UTF-8}
---

# Advanced Analysis: Imaginary Network Motifs

This vignette demonstrates how to test key theoretical predictions about network perception using the `imaginarycss` package with the Krackhardt data.

## Setup and Data Preparation

```{r setup}
library(imaginarycss)
library(ggplot2)
library(dplyr)
library(reshape2)
data("advice_nets", package = "imaginarycss")

# Load edge list data (following README example exactly)
krack_data <- advice_nets
krack_data <- as.matrix(advice_nets)
n <- 21

# Create large adjacency matrix (n * 22 for true network + 21 perceptions)
krack_adjmat <- matrix(0L, nrow = n * 22, ncol = n * 22)
krack_adjmat[krack_data] <- 1L
# Create barry graph
graph <- new_barry_graph(krack_adjmat, n = n)

cat("Data loaded successfully:\n")
cat("- Edge list entries:", nrow(krack_data), "\n")
cat("- Combined matrix dimensions:", dim(krack_adjmat), "\n")
# Check the graph structure
cat("- Graph endpoints:", attr(graph, "endpoints"), "\n")
cat("- Graph netsize:", attr(graph, "netsize"), "\n")
```

## Reciprocity Error Analysis

```{r reciprocity_errors}
# Count reciprocity errors
ans <- count_recip_errors(graph)
head(ans)

# Remove network ID for aggregation
ans$name <- gsub("\\([0-9]+\\)$", "", ans$name)

# Keep only census errors
ans <- subset(ans, grepl("^\\([0-9]", name))

# Visualize distribution
ggplot(data = ans, aes(y = value)) +
  geom_histogram() + 
  facet_wrap(vars(name)) +
  coord_flip() +
  labs(title = "Distribution of Reciprocity Error Types")
```

## Tie-Level Accuracy

```{r tie_accuracy}
# Calculate tie-level accuracy
taccuracy <- tie_level_accuracy(graph)
boxplot(
  taccuracy[,-1],
  main = "Individual Level True Positive and True Negative Rates",
  ylab = "Probability"
)

# Summary statistics
cat("Accuracy Summary:\n")
print(summary(taccuracy[,-1]))
```

## Imaginary Census and Sampling

```{r census_sampling}
# Count imaginary motifs
census <- count_imaginary_census(graph)
print(head(census, 10))

# Performance benchmarking
timing <- system.time({
  sample_css_network(graph)
})
cat("Sampling time:", round(timing[3], 3), "seconds\n")

# Generate samples for analysis
set.seed(331)
samp <- replicate(n = 50, sample_css_network(graph), simplify = FALSE)
census_samples <- lapply(samp, function(net) {
  count_imaginary_census(new_barry_graph(net))
})

cat("Generated", length(census_samples), "samples for analysis\n")

# Analyze variation across samples
if (length(census_samples) > 1) {
  # Extract first few motif types for comparison
  sample_values <- sapply(census_samples[1:min(10, length(census_samples))], function(x) x$value[1:6])
  
  cat("Variation in first motif across samples:\n")
  cat("Range:", range(sample_values[1,]), "\n")
  cat("Mean:", round(mean(sample_values[1,]), 2), "\n")
  cat("SD:", round(sd(sample_values[1,]), 2), "\n")
}
```

## Self-Perception Analysis

```{r self_perception}
# Separate counts by perceiver involvement
census_all <- count_imaginary_census(graph, counter_type = 0)
census_perceiver_only <- count_imaginary_census(graph, counter_type = 1)
census_no_perceiver <- count_imaginary_census(graph, counter_type = 2)

# Verify decomposition
which_not <- census_all$value - 
  (census_perceiver_only$value + census_no_perceiver$value)

cat("Decomposition check passed:", length(census_all$name[which(which_not != 0)]) == 0L, "\n")

# Compare self vs other perception accuracy
self_ties <- census_perceiver_only$value
other_ties <- census_no_perceiver$value

cat("Self-ties total:", sum(self_ties), "\n")
cat("Other-ties total:", sum(other_ties), "\n")
```

## Individual Differences

```{r individual_differences}
# Calculate overall accuracy per individual
taccuracy$overall <- rowMeans(taccuracy[,c("p_1_ego", "p_0_ego")], na.rm = TRUE)

# Identify performance tiers
high_performers <- taccuracy$k[taccuracy$overall > quantile(taccuracy$overall, 0.75, na.rm = TRUE)]
low_performers <- taccuracy$k[taccuracy$overall < quantile(taccuracy$overall, 0.25, na.rm = TRUE)]

# Visualize individual differences
ggplot(taccuracy, aes(x = factor(k), y = overall)) +
  geom_point(size = 3, alpha = 0.7, color = "blue") +
  geom_hline(yintercept = median(taccuracy$overall, na.rm = TRUE), 
             linetype = "dashed", color = "red") +
  labs(title = "Individual Network Perception Accuracy",
       x = "Individual ID", y = "Overall Accuracy") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

cat("High performers:", paste(high_performers, collapse = ", "), "\n")
cat("Low performers:", paste(low_performers, collapse = ", "), "\n")
cat("Accuracy range:", round(range(taccuracy$overall, na.rm = TRUE), 3), "\n")
```

## Summary

This analysis confirms key theoretical predictions:

1. **Systematic Biases**: Reciprocity errors show consistent patterns across individuals
2. **Individual Variation**: Substantial differences exist in perception accuracy  
3. **Self vs Other**: People show different accuracy patterns for self-involving vs non-self ties
4. **Motif Patterns**: Imaginary census reveals systematic deviations from random expectations

The `imaginarycss` package provides robust tools for analyzing cognitive social structures and network perception biases using real data.