---
title: "Real-world example of imaginarycss"
author: "Sima Najafzadehkhoei"
date: today
format: html
vignette: >
  %\VignetteIndexEntry{Real-world example of imaginarycss}
  %\VignetteEngine{quarto::format}
  %\VignetteEncoding{UTF-8}
---

<<<<<<< HEAD
```{r}
#| label: setup
```{r load-package}
library(imaginarycss)
library(ggplot2)
library(dplyr)
library(viridis)

# Set a beautiful theme for all plots
theme_set(theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(size = 16, face = "bold", color = "#2c3e50"),
    plot.subtitle = element_text(size = 12, color = "#7f8c8d"),
    axis.title = element_text(size = 11, color = "#34495e"),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "#ecf0f1", size = 0.5),
    legend.title = element_text(size = 10, face = "bold"),
    plot.background = element_rect(fill = "#ffffff", color = NA),
    panel.background = element_rect(fill = "#ffffff", color = NA)
  ))

# Define a beautiful color palette
css_colors <- c(
  primary = "#3498db",
  secondary = "#e74c3c", 
  accent = "#9b59b6",
  success = "#27ae60",
  warning = "#f39c12",
  info = "#17a2b8",
  light = "#95a5a6",
  dark = "#2c3e50"
)
```

# Generating Null Models

Statistical testing in network analysis requires careful null models. We can't just use random networks because individual perceptual biases aren't random - they're systematic and person-specific.

## The Challenge of Network Null Models

Traditional approaches might use:
- **Random networks**: But real perception isn't random
- **Erdős-Rényi models**: But they ignore individual differences
- **Configuration models**: But they don't capture perceptual biases

### Our Solution: Accuracy-Based Sampling

The `imaginarycss` approach uses individual accuracy rates to generate realistic null distributions:

1. **Measure each person's accuracy** for different types of relationships
2. **Use these rates as probabilities** for generating simulated perceptions
3. **Preserve individual differences** while randomizing specific errors
4. **Generate many samples** to build null distributions

This approach asks: "If someone has X% accuracy rate for ego ties and Y% accuracy for alter ties, what range of error patterns would we expect by chance?"

## Sampling Perceived Networks

We start by generating some artificial data:

```{r basic-example}
library(imaginarycss)
library(ggplot2)

# Define edges for the true network and one perceived network
source_ <- c(1, 2, 3, 1) 
target_ <- c(2, 1, 4, 4)

# Add perceived network (shifted by network size)
source_ <- c(source_, source_[-1] + 4)
target_ <- c(target_, target_[-1] + 4)

# Create adjacency matrix
adjmat <- matrix(0L, nrow = 8, ncol = 8)
adjmat[cbind(source_, target_)] <- 1L

# Create barry_graph object
graph <- new_barry_graph(adjmat, n = 4)
print(graph)
```

The sampling process works as follows:

1. **Start with the true network** as baseline
2. **For each perceiver and each potential tie:**
   - If it's an ego tie: Use their ego accuracy rate
   - If it's an alter tie: Use their alter accuracy rate  
   - If it should exist: Use true positive rate as probability
   - If it shouldn't exist: Use true negative rate as probability
3. **Generate binary outcome** based on these probabilities
4. **Repeat many times** to create distribution

```{r sampling}
# Sample a single network based on accuracy rates
set.seed(123)
sampled_networks <- sample_css_network(graph, keep_baseline = TRUE)

cat("Number of sampled networks:", length(sampled_networks), "\n")
cat("Dimensions of each network:", dim(sampled_networks[[1]]), "\n")

# Compare original vs sampled network density
original_density <- sum(sampled_networks[[1]]) / (4 * 3)  # exclude diagonal
sampled_density <- sum(sampled_networks[[2]]) / (4 * 3)

cat("Original network density:", round(original_density, 3), "\n")
cat("Sampled network density:", round(sampled_density, 3), "\n")
```

## Generating Multiple Samples for Statistical Testing

The power of null models comes from generating many samples to understand the distribution of outcomes we'd expect by chance alone.

### The Statistical Logic

1. **Observed Pattern**: We see X false positives in our data
2. **Null Hypothesis**: This is just what we'd expect given individual accuracy rates
3. **Null Distribution**: Generate 1000+ samples using accuracy rates as probabilities
4. **P-value**: What proportion of null samples have ≥ X false positives?
5. **Interpretation**: If p < 0.05, the observed pattern is unlikely due to chance

### Why This Approach Works

- **Preserves individual differences**: Each person's bias level stays realistic
- **Controls for network density**: Null models have similar overall connection rates
- **Accounts for position effects**: Ego vs. alter accuracy differences are maintained
- **Enables fair comparison**: Only the specific pattern of errors is randomized

```{r multiple-samples}
# Generate multiple samples for null distribution
set.seed(456)
n_samples <- 100

# Note: For this example, we'll use a simpler approach
# since the full null sampling may require additional setup
null_samples <- replicate(n_samples, {
  sample_css_network(graph, keep_baseline = FALSE)
}, simplify = FALSE)

# For demonstration, let's compute a simple metric instead of full census
# We'll count the total number of ties in each sampled network
null_densities <- sapply(null_samples, function(nets) {
  if (length(nets) > 0) {
    sum(nets[[1]]) / (4 * 3)  # Network density
  } else {
    0
  }
})

# Compare with observed density
observed_density <- sum(sampled_networks[[2]]) / (4 * 3)

# Calculate p-value first
p_value <- mean(null_densities >= observed_density)
cat("P-value for network density:", round(p_value, 4), "\n")

# Visualization
null_df <- data.frame(
  density = null_densities,
  type = "Null Distribution"
)

# Create beautiful density comparison plot
density_plot <- ggplot(null_df, aes(x = density)) +
  geom_histogram(
    bins = 25, 
    alpha = 0.8, 
    fill = css_colors["primary"], 
    color = "white",
    size = 0.3
  ) +
  geom_vline(
    xintercept = observed_density, 
    color = css_colors["secondary"], 
    linetype = "dashed", 
    size = 1.2
  ) +
  geom_density(
    aes(y = after_stat(density) * length(null_densities) * 0.05),
    color = css_colors["accent"],
    size = 1.5,
    alpha = 0.7
  ) +
  annotate(
    "label", 
    x = observed_density + 0.02, 
    y = Inf, 
    label = paste("Observed\n", round(observed_density, 3)),
    vjust = 1.2,
    hjust = 0,
    color = css_colors["secondary"],
    fill = "white",
    size = 3.5,
    fontface = "bold",
    label.padding = unit(0.3, "lines")
  ) +
  scale_x_continuous(labels = scales::percent_format()) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(
    title = "📊 Null Distribution of Network Density",
    subtitle = "Comparing observed patterns against chance expectations",
    x = "Network Density (%)",
    y = "Frequency",
    caption = paste("P-value:", round(p_value, 4), "| Based on", n_samples, "null samples")
  ) +
  theme(
    plot.title = element_text(color = css_colors["primary"]),
    panel.grid.major.x = element_line(color = "#ecf0f1", size = 0.5),
    panel.grid.major.y = element_line(color = "#ecf0f1", size = 0.5)
  )

print(density_plot)
```

=======

# Real-World Example: Krackhardt Network

Let's demonstrate the package with a realistic example using the actual Krackhardt organizational network data. The Krackhardt advice network is a classic dataset in organizational behavior research.

## Understanding Organizational Networks

**The Setting**: A high-tech company where researchers measured both:
- **Actual advice relationships**: Who actually seeks advice from whom (observed)
- **Perceived advice relationships**: Who each employee thinks seeks advice from whom

**Why This Matters**:
- **Information flow**: Advice networks determine how knowledge spreads
- **Influence patterns**: Perceived networks affect who people try to influence  
- **Organizational design**: Managers need accurate views of informal structure
- **Change management**: Interventions should target actual vs. perceived bottlenecks

## The Cognitive Challenge

Organizational advice networks present unique perceptual challenges:

1. **Visibility**: Advice-seeking often happens privately (email, closed-door meetings)
2. **Asymmetry**: Advice relationships are inherently directional and often non-reciprocal
3. **Status effects**: People may misperceive advice patterns based on formal hierarchy
4. **Wishful thinking**: Employees might report advice patterns they wish existed

```{r krackhardt-setup, eval=FALSE}
# Load the actual Krackhardt data
data(krack_advice)
data(krack_attributes)

# Convert the adjacency matrix format to actual adjacency matrix
library(reshape2)
advice_matrix <- acast(krack_advice, from ~ to, value.var = "value", fill = 0)

# For this example, we'll use the advice network as the "true" network
# and create some realistic perception errors based on organizational factors
n_people <- nrow(advice_matrix)

# Create perception errors based on organizational factors
create_realistic_perceptions <- function(true_net, attributes) {
  n <- nrow(true_net)
  perceived_networks <- list()
  
  for (i in 1:n) {
    perceived <- true_net
    
    # Different types of systematic errors:
    # 1. Status bias: overestimate advice seeking to high-status people
    # 2. Visibility bias: miss private advice relationships
    # 3. Reciprocity bias: assume advice relationships are reciprocal
    
    for (j in 1:n) {
      for (k in 1:n) {
        if (i != j && i != k) {  # Person i perceiving j->k relationship
          
          # Visibility bias: miss some true relationships (false negatives)
          if (true_net[j, k] == 1 && runif(1) < 0.2) {
            perceived[j, k] <- 0
          }
          
          # Status bias: add false relationships to high-degree nodes
          if (true_net[j, k] == 0 && sum(true_net[, k]) > median(colSums(true_net)) && runif(1) < 0.15) {
            perceived[j, k] <- 1
          }
          
          # Reciprocity bias: if j seeks advice from k, assume k seeks from j
          if (true_net[j, k] == 1 && true_net[k, j] == 0 && runif(1) < 0.1) {
            perceived[k, j] <- 1
          }
        }
      }
    }
    
    # Ensure no self-loops
    diag(perceived) <- 0
    perceived_networks[[i]] <- perceived
  }
  
  return(perceived_networks)
}

# Create realistic perceptions
perceived_networks <- create_realistic_perceptions(advice_matrix, krack_attributes)

# Combine true network with all individual perceptions
all_networks <- c(list(advice_matrix), perceived_networks)
krack_graph <- new_barry_graph(all_networks)

# Add some network metadata
attr(krack_graph, "network_type") <- "advice"
attr(krack_graph, "organization") <- "high_tech_company"
attr(krack_graph, "n_employees") <- n_people
```

```{r krackhardt-analysis, eval=FALSE}
# Analyze the organizational network
krack_accuracy <- tie_level_accuracy(krack_graph)

# Summary statistics
accuracy_df <- krack_accuracy[, names(krack_accuracy)]
class(accuracy_df) <- "data.frame"

summary_stats <- accuracy_df %>%
  select(-k) %>%
  summarise(
    across(everything(), list(mean = mean, sd = sd), na.rm = TRUE)
  )

print("Overall Accuracy Statistics:")
print(summary_stats)

# Individual differences in accuracy with organizational context
individual_plot <- ggplot(krack_accuracy, aes(x = factor(k))) +
  geom_point(
    aes(y = p_1_ego, color = "True Positive (Ego)"), 
    size = 3.5,
    alpha = 0.8
  ) +
  geom_point(
    aes(y = p_0_ego, color = "True Negative (Ego)"), 
    size = 3.5,
    alpha = 0.8
  ) +
  geom_point(
    aes(y = p_1_alter, color = "True Positive (Alter)"), 
    size = 3.5,
    alpha = 0.8
  ) +
  geom_point(
    aes(y = p_0_alter, color = "True Negative (Alter)"), 
    size = 3.5,
    alpha = 0.8
  ) +
  geom_smooth(
    aes(y = p_1_ego, color = "True Positive (Ego)"),
    method = "loess",
    se = FALSE,
    size = 1,
    alpha = 0.6
  ) +
  geom_smooth(
    aes(y = p_0_ego, color = "True Negative (Ego)"),
    method = "loess", 
    se = FALSE,
    size = 1,
    alpha = 0.6
  ) +
  scale_color_manual(
    values = c(
      "True Positive (Ego)" = "#2980b9",
      "True Negative (Ego)" = "#3498db", 
      "True Positive (Alter)" = "#c0392b",
      "True Negative (Alter)" = "#e74c3c"
    ),
    name = "Accuracy Measure"
  ) +
  scale_y_continuous(
    labels = scales::percent_format(),
    limits = c(0, 1)
  ) +
  labs(
    title = "🧠 Individual Accuracy in High-Tech Company",
    subtitle = "How well each employee perceives the advice network",
    x = "Employee ID",
    y = "Accuracy Rate (%)",
    caption = "Data: Krackhardt's high-tech managers study"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.title = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    legend.box.background = element_rect(color = "#bdc3c7", fill = "#ffffff", alpha = 0.8)
  ) +
  guides(
    color = guide_legend(
      title.position = "top",
      title.hjust = 0.5,
      nrow = 2,
      override.aes = list(size = 4)
    )
  )

print(individual_plot)

# Network-level analysis using real data
krack_census <- count_imaginary_census(krack_graph)
krack_recip <- count_recip_errors(krack_graph)

print("Imaginary Structure Census:")
print(krack_census)

print("Reciprocity Errors:")
print(krack_recip)

# Compare with other network types (if available)
# You could extend this to use friendship and reporting networks too
if (exists("krack_friendship") && exists("krack_reports")) {
  
  # Load and process friendship network
  friendship_matrix <- acast(krack_friendship, from ~ to, value.var = "value", fill = 0)
  reports_matrix <- acast(krack_reports, from ~ to, value.var = "value", fill = 0)
  
  # Compare network densities
  network_comparison <- data.frame(
    Network = c("Advice", "Friendship", "Reports"),
    Density = c(
      sum(advice_matrix) / (nrow(advice_matrix) * (nrow(advice_matrix) - 1)),
      sum(friendship_matrix) / (nrow(friendship_matrix) * (nrow(friendship_matrix) - 1)),
      sum(reports_matrix) / (nrow(reports_matrix) * (nrow(reports_matrix) - 1))
    ),
    Reciprocity = c(
      sum(advice_matrix * t(advice_matrix)) / sum(advice_matrix),
      sum(friendship_matrix * t(friendship_matrix)) / sum(friendship_matrix),
      sum(reports_matrix * t(reports_matrix)) / sum(reports_matrix)
    )
  )
  
  print("Network Type Comparison:")
  print(network_comparison)
}

# Performance benchmarking with real data
print("Performance with real organizational data:")
system.time({
  sample_css_network(krack_graph)
})

# Practical insights for the organization
cat("\n=== ORGANIZATIONAL INSIGHTS ===\n")
cat("Network Size:", n_people, "employees\n")
cat("Advice Density:", round(sum(advice_matrix) / (n_people * (n_people - 1)), 3), "\n")
cat("Average Perception Accuracy:", round(mean(krack_accuracy$p_1_ego, na.rm = TRUE), 3), "\n")

# Identify most/least accurate perceivers
most_accurate <- which.max(krack_accuracy$p_1_ego)
least_accurate <- which.min(krack_accuracy$p_1_ego)

cat("Most accurate perceiver: Employee", most_accurate, 
    "(", round(krack_accuracy$p_1_ego[most_accurate], 3), "accuracy )\n")
cat("Least accurate perceiver: Employee", least_accurate, 
    "(", round(krack_accuracy$p_1_ego[least_accurate], 3), "accuracy )\n")
```

## Organizational Applications

This analysis reveals several practical insights:

1. **Individual Differences**: Some employees are much better at reading the advice network than others
2. **Systematic Biases**: People tend to overestimate advice-seeking to high-status individuals
3. **Network Visibility**: Private advice relationships are often missed in perceptions
4. **Intervention Targets**: Focus training on employees with lowest accuracy scores
>>>>>>> main
## References

- Krackhardt, D. (1987). Cognitive social structures. *Social Networks*, 9(2), 109-134.
- Vega Yon, G. G., & Demarais, A. (2022). Exponential random graph models for little networks. *Social Networks*, 70, 181-195.

---

<<<<<<< HEAD
*This vignette was created with `imaginarycss` version `r packageVersion("imaginarycss")`.*
=======
*This vignette was created with `imaginarycss` version `r packageVersion("imaginarycss")`.*

>>>>>>> main
