---
title: "BNAM & ERGM - Imaginary Network Motifs"
output: html_document
date: "`r Sys.Date()`"
---

This script runs BNAM with graph mixture prior (Butt, 2014) and compared it with perceived networks.

Read the packages and source the BNAM extension file.

```{r setup, message = FALSE, warning = FALSE}
# Read packages
library(statnet) # read the bundle of statnet packages
library(tidyverse) # general data processing
library(imaginarycss) # css manipulation
library(ergmito) # George's ERGM package
# library(netdiffuseR)
library(netplot) # network visualization
library(ggplot2) # general visualization
library(ggpubr) # add stat to ggplot
library(grid) # add grid to ggplot
library(gridExtra) # add some notations to ggplot
library(xtable) # create a latex table
library(rstatix)

# Load the CSS data sets
load('../data/replication1.RData')

# Load BNAM mixture code
source('../R/mixture.code.R')
```

We use 8 data sets listed below for this study:

1. Silicon System Advice Seeking
2. Silicon Systems Friendship
3. High Tech Managers Advice Seeking
4. High Tech Managers Friendship
5. Italian University Advice Seeking
6. Italian University Friendship
7. Pacific Distributors Advice Seeking
8. Pacific Distributors Friendship

```{r BNAM}
# Function to run BNAM with graph mixture prior on a CSS data set
css_bbnam_mix <- function(css_array, # input CSS array
                          emp = c(1, 11), # input priors for false negatives
                          epp = c(1, 11)){ # input priors for false positives
  # Extract the size of css_array
  mat_size <- nrow(css_array)
  # np <- matrix(0.5,mat_size,mat_size)  # 21 x 21 matrix of Bernoulli parameters (since n=21)
  # Run graph mixture prior model
  mix.actor <- bbnam.mix(lapply(1:mat_size,function(x) as.network(css_array[x,,])),
                          nprior = NULL, emprior = c(1, 11), epprior = c(1, 11), 
                          diag = FALSE, mode = "digraph", reps = 5, draws = 1500, 
                          burntime = 500, quiet = TRUE, anames = NULL, onames = NULL, 
                          compute.sqrtrhat = TRUE)
  # Save it as a network object
  mix.net <- as.network(apply(mix.actor$net, c(2, 3), median))
  return(mix.net)
}

# save it as another obj
# .true is based on BNAM 
# ital_ad.true <- ad.true 
# ital_fr.true <- fr.true
# .mix is based on BNAM with graph mixture prior
ital_ad.mix <- css_bbnam_mix(ital_ad)
ital_fr.mix <- css_bbnam_mix(ital_fr)
# sili_ad.true <- ad.true
# sili_fr.true <- fr.true
sili_ad.mix <- css_bbnam_mix(sili_ad)
sili_fr.mix <- css_bbnam_mix(sili_fr)
# high_ad.true <- ad.true
# high_fr.true <- fr.true
high_ad.mix <- css_bbnam_mix(high_ad)
high_fr.mix <- css_bbnam_mix(high_fr)
# paci_ad.true <- ad.true
# paci_fr.true <- fr.true
paci_ad.mix <- css_bbnam_mix(paci_ad)
paci_fr.mix <- css_bbnam_mix(paci_fr)
```

Now, we look at some descriptive statistics of these networks.

```{r descriptive stats}
# Function to create a descriptive stat table
desc_table <- function(net_name = c('HM AFN','HM PFN', 'HM AANs','HM PANs'),
                       fr.mix = high_fr.mix,
                       ad.mix = high_ad.mix,
                       css_fr = high_fr,
                       css_ad = high_ad){
  mat_size <-  nrow(css_fr)
  table <- data.frame(Network = net_name,
                    People = c(network.size(fr.mix),
                               paste(round(mean(sapply(lapply(1:mat_size,function(x) as.network(css_fr[x,,])),network.size)),2),'\u00B1',
                                     round(sd(sapply(lapply(1:mat_size,function(x) as.network(css_fr[x,,])),network.size)),2)),
                               network.size(ad.mix),
                               paste(round(mean(sapply(lapply(1:mat_size,function(x) as.network(css_ad[x,,])),network.size)),2),'\u00B1',
                                     round(sd(sapply(lapply(1:mat_size,function(x) as.network(css_ad[x,,])),network.size)),2))),
                    Edges = c(network.edgecount(fr.mix),
                              paste(round(mean(sapply(lapply(1:mat_size,function(x) as.network(css_fr[x,,])),network.edgecount)),2),'\u00B1',
                                    round(sd(sapply(lapply(1:mat_size,function(x) as.network(css_fr[x,,])),network.edgecount)),2)),
                              network.edgecount(ad.mix),
                              paste(round(mean(sapply(lapply(1:mat_size,function(x) as.network(css_ad[x,,])),network.edgecount)),2),'\u00B1',
                                    round(sd(sapply(lapply(1:mat_size,function(x) as.network(css_ad[x,,])),network.edgecount)),2))),
                    Density = c(round(network.density(fr.mix),2),
                                paste(round(mean(sapply(lapply(1:mat_size,function(x) as.network(css_fr[x,,])),network.density)),2),'\u00B1',
                                      round(sd(sapply(lapply(1:mat_size,function(x) as.network(css_fr[x,,])),network.density)),2)),
                                round(network.density(ad.mix),2),
                                paste(round(mean(sapply(lapply(1:mat_size,function(x) as.network(css_ad[x,,])),network.density)),2),'\u00B1',
                                      round(sd(sapply(lapply(1:mat_size,function(x) as.network(css_ad[x,,])),network.density)),2))),
                    Mutual = c(dyad.census(fr.mix)[[1,1]],
                               paste(round(mean(sapply(lapply(1:mat_size,function(x) as.network(css_fr[x,,])), function(x) dyad.census(x)[[1,1]])),2),'\u00B1',
                                     round(sd(sapply(lapply(1:mat_size,function(x) as.network(css_fr[x,,])), function(x) dyad.census(x)[[1,1]])),2)),dyad.census(ad.mix)[[1,1]],
                               paste(round(mean(sapply(lapply(1:mat_size,function(x) as.network(css_ad[x,,])), function(x) dyad.census(x)[[1,1]])),2),'\u00B1',
                                     round(sd(sapply(lapply(1:mat_size,function(x) as.network(css_ad[x,,])), function(x) dyad.census(x)[[1,1]])),2))),
                    Asymmetric = c(dyad.census(fr.mix)[[1,2]],
                                   paste(round(mean(sapply(lapply(1:mat_size,function(x) as.network(css_fr[x,,])), function(x) dyad.census(x)[[1,2]])),2),'\u00B1',
                                         round(sd(sapply(lapply(1:mat_size,function(x) as.network(css_fr[x,,])), function(x) dyad.census(x)[[1,2]])),2)),dyad.census(ad.mix)[[1,2]],
                                   paste(round(mean(sapply(lapply(1:mat_size,function(x) as.network(css_ad[x,,])), function(x) dyad.census(x)[[1,2]])),2),'\u00B1',
                                         round(sd(sapply(lapply(1:mat_size,function(x) as.network(css_ad[x,,])), function(x) dyad.census(x)[[1,2]])),2))),
                    Null = c(dyad.census(fr.mix)[[1,3]],
                             paste(round(mean(sapply(lapply(1:mat_size,function(x) as.network(css_fr[x,,])), function(x) dyad.census(x)[[1,3]])),2),'\u00B1',
                                   round(sd(sapply(lapply(1:mat_size,function(x) as.network(css_fr[x,,])), function(x) dyad.census(x)[[1,3]])),2)),dyad.census(ad.mix)[[1,3]],
                             paste(round(mean(sapply(lapply(1:mat_size,function(x) as.network(css_ad[x,,])), function(x) dyad.census(x)[[1,3]])),2),'\u00B1',
                                   round(sd(sapply(lapply(1:mat_size,function(x) as.network(css_ad[x,,])), function(x) dyad.census(x)[[1,3]])),2))))
  return(table)
}

table <- do.call('rbind',list(desc_table(net_name = c('SS AFN','SS PFN', 'SS AANs','SS PANs'),sili_fr.mix,sili_ad.mix,sili_fr,sili_ad),
                              desc_table(net_name = c('HM AFN','HM PFN', 'HM AANs','HM PANs'),high_fr.mix,high_ad.mix,high_fr,high_ad),
                              desc_table(net_name = c('IU AFN','IU PFN', 'IU AANs','IU PANs'),ital_fr.mix,ital_ad.mix,ital_fr,ital_ad),
                              desc_table(net_name = c('PD AFN','PD PFN', 'PD AANs','PD PANs'),paci_fr.mix,paci_ad.mix,paci_fr,paci_ad)))
xtable(table,include.rownames = FALSE)
```

