---
title: "Illustrative Examples of Imaginary Structures"
output:
  html_document: default
  pdf_document: default
editor_options: 
  chunk_output_type: console
---
This doc shows some illustrative examples of imaginary structures in CSS data.

## Glossary
- __P-EO__: Partially false errors of omission
- __P-EC__: Partially false errors of commission
- __C-EO__: Completely false errors of omission
- __C-EC__: Completely false errors of commission
- __M-E__: Mixed false errors

## Installation
To calculate imaginary structure statistics, you can install the `imaginarycss` package from [CRAN](https://CRAN.R-project.org).

```r
# install.packages("imaginarycss")
```
Or you can also install the development version from [GitHub](https://github.com/).
``` r
# install.packages("devtools")
# devtools::install_github("gvegayon/imaginary-structures/imaginarycss@main")
```

## 3-node Graph Case with 2-node Imaginary Motifs
First, we define a true graph. In this case, the true graph includes 3 nodes and 3 edges. It also has all the possible dyadic MAN conditions:

- mutual
- asymmetric
- null

Consequently, this true graph example can cover all the possible dyadic imaginary motifs, compared to different perceived graph patterns.

In the following, we're also going to construct three perceived graphs that compare with the true graph. we calculate example imaginary motif statistics using these examples.

```{r message = FALSE, warning = FALSE}
library(sna)
library(imaginarycss)
library(ergmito)
set.seed(1231)
```

```{r true graph}
# library(imaginarycss)
# 3 actor true graph
t_source_ <- c("A","A","B") 
t_target_ <- c("B","C","A")
t_source <- c(1,1,2)
t_target <- c(2,3,1)
t_ <- as.network(data.frame(from = t_source_,
                      to = t_target_))

coords <- gplot(
  t_, displaylabels = T, usecurve = T, edge.curve = .1,
  main = "True", jitter = FALSE
  )
```

### A's perceived graph
We define A's perceived graph and then compare it with the true graph. By doing so, we should obverse the following imaginary 2-node motifs:

- P-EO = 1
- P-EC = 1
- C-EO = 0
- C-EC = 0


```{r A graph}
# A's perceived graph
a_source_ <- c("A","A","C")
a_target_ <- c("B","C","A")
a_source <- c(1,1,3)
a_target <- c(2,3,1)
a_ <- as.network(data.frame(from = a_source_,
                 to = a_target_))
gplot(
  a_, displaylabels = T, usecurve = T, edge.curve = .1,
  coord = coords, main = "A", jitter = FALSE
  )
```

### B's perceived graph
Compared to the true graph, B should provide the following imaginary 2-node motifs:

- P-EO = 0
- P-EC = 0
- C-EO = 2
- C-EC = 3

```{r B graph}
# C's perceived graph
b_source_ <- c("C","B","C")
b_target_ <- c("A","C","B")
b_source <- c(3,2,3) -1
b_target <- c(1,3,2) -1
b_ <- as.network(data.frame(from = b_source_,
                            to = b_target_),matrix.type='edgelist')
gplot(
  b_, displaylabels = T, usecurve = T, edge.curve = .1,
  coord = coords, main = "B", jitter = FALSE
  )
```

### C's perceived graph
This should be C's perceived graph statistics:

- P-EO = 2
- P-EC = 1
- C-EO = 0
- C-EC = 0

```{r C graph}
# C's perceived graph
c_source_ <- c("C","B")
c_target_ <- c("A","A")
c_source <- c(3,2) -1
c_target <- c(1,1) -1
c_ <- as.network(data.frame(from = c_source_,
                      to = c_target_),matrix.type='edgelist')
gplot(
  c_, displaylabels = T, usecurve = T, edge.curve = .1,
  coord = coords, main = "C", jitter = FALSE
  )
```

### Count Imaginary Errors

```{r all-three, echo = FALSE}
op <- par(mfcol=c(2, 2), mar = rep(1, 4))
gplot(
  t_, displaylabels = T, usecurve = T, edge.curve = .1,
  coord = coords, main = "True", jitter = FALSE
  )
gplot(
  a_, displaylabels = T, usecurve = T, edge.curve = .1,
  coord = coords, main = "A", jitter = FALSE)

gplot(
  b_, displaylabels = T, usecurve = T, edge.curve = .1,
  coord = coords, main = "B", jitter = FALSE)

gplot(
  c_, displaylabels = T, usecurve = T, edge.curve = .1,
  coord = coords, main = "C", jitter = FALSE)
par(op)
```


Let's count errors using `count_recip_errors` in the imaginary_css package.

```{r Errors}
big_graph <- blockdiagonalize(list(t_, a_, b_, c_))
graph <- new_barry_graph(as.matrix(big_graph), n = 3)

ans <- count_recip_errors(graph)
ans$name <- gsub("[(][0-9]+[)]", "", ans$name)
ans
```

<!-- > ans -->
<!--   id                                name value -->
<!-- 1  0   Partially false recip (omission)      1 -->
<!-- 2  1   Partially false recip (omission)      0 -->
<!-- 3  0  Partially false recip (comission)      1 -->
<!-- 4  1  Partially false recip (comission)      0 -->
<!-- 5  0  Completely false recip (omission)      0 -->
<!-- 6  1  Completely false recip (omission)      1 -->
<!-- 7  0 Completely false recip (comission)      0 -->
<!-- 8  1 Completely false recip (comission)      1 -->


## Krackhardt's CSS data

We have access to a CSS data set that Krackhardt (1987) collected.

```{r Krackhardt CSS}
krack_css <- load('../data/krackhardt_css.RData')

# the case of friendship
true <- as.network(consensus(friendship_nets,mode="digraph",diag=FALSE,method="OR.row"))
# true <- as.network(apply(fr.post.actor$net, c(2, 3), median)) #bbnam as true
big_krack <- blockdiagonalize(c(list(true),friendship_nets))
# the case of advice
true <- as.network(consensus(advice_nets,mode="digraph",diag=FALSE,method="OR.row"))
big_krack <- blockdiagonalize(c(list(true),advice_nets))

# create a barry graph object
krack_graph <- new_barry_graph(as.matrix(big_krack), n = 21)
ans <- count_recip_errors(krack_graph)
ans$name <- gsub("[(][0-9]+[)]", "", ans$name)
ans
```

BNAM (Baysian network inference): Let's start by defining some priors for the Bayesian network inference model. We'll use an uninformative network prior, together with weakly informative (but diffuse and symmetric) priors on the error rates. Read the man page ("?bbnam") to get more information about how the routine works.

Check Lee & Butts (2018; 2020) to get their parameters.
```{r BNAM}
np<-matrix(0.5,21,21)  # 21 x 21 matrix of Bernoulli parameters (since n=21)
emp<-sapply(c(1,11),rep,21)  # Beta(3,11) priors for false negatives
epp<-sapply(c(1,11),rep,21)  # Beta(3,11) priors for false positives
hist(rbeta(100000,3,11))  # This gives you a sense of what the priors look like!

# Now, let's take some posterior draws for the friendship network, using 
# various models (warning: slow)
# fr.post.fixed<-bbnam.fixed(friendship_nets,nprior=np,em=3/(3+11),ep=3/(3+11))
# fr.post.pooled<-bbnam.pooled(friendship_nets,nprior=np,em=emp[1,],ep=epp[1,])
fr.post.actor<-bbnam.actor(friendship_nets,nprior=np,em=emp,ep=epp)
summary(fr.post.actor)
# plot(fr.post.actor)
gplot(as.network(apply(fr.post.actor$net, c(2, 3), median)))

# advice
ad.post.actor<-bbnam.actor(advice_nets,nprior=np,em=emp,ep=epp)
gplot(as.network(apply(ad.post.actor$net, c(2, 3), median)))
```
```{r network plot}
library(netplot)
library(gridExtra)
g_ad <- igraph::graph_from_adjacency_matrix(as.matrix(ad.true))
g_fr <- igraph::graph_from_adjacency_matrix(as.matrix(fr.true))
ad_1 <- advice_nets[[1]]
fr_1 <- friendship_nets[[1]]
fr_fp <- igraph::graph_from_adjacency_matrix(
  ifelse( (as.matrix(fr.true) - as.matrix(fr_1)) == -1,1,0)
)
fr_fn <- igraph::graph_from_adjacency_matrix(
  ifelse( (as.matrix(fr.true) - as.matrix(fr_1)) == 1,1,0)
)
ad_fp <- igraph::graph_from_adjacency_matrix(
  ifelse( (as.matrix(ad.true) - as.matrix(ad_1)) == -1,1,0)
)
ad_fn <- igraph::graph_from_adjacency_matrix(
  ifelse( (as.matrix(ad.true) - as.matrix(ad_1)) == 1,1,0)
)
layout <- igraph::layout_with_dh(g_ad)
# png(file = 'netplot_v1.png')
gviz <- grid.arrange(
  arrangeGrob(nplot(
        g_fr,
        layout = layout,
        vertex.label = NULL,
        dge.curvature = pi/4,
        edge.width = .8,
        edge.col = ~ego(alpha=0, col = '#e88e2d') + alter(alpha = .7)), left = 'Friendship Network', top='Actual'),
  arrangeGrob(nplot(fr_1,
        layout = layout,
        vertex.label = NULL,
        edge.curvature = pi/4,
        edge.line.lty = 3,
        edge.width = .8,
        edge.col = ~ego(alpha=0, col = '#e88e2d') + alter(alpha = .7)),top='Perceived by ID 1'),
  arrangeGrob(nplot(fr_fp,
        layout = layout,
        vertex.label = NULL,
        edge.curvature = pi/4,
        edge.width = .8,
        edge.col = ~ego(alpha=0, col = '#55B748') + alter(alpha = .7)),top='False Positive'),
  arrangeGrob(nplot(fr_fn,
        layout = layout,
        vertex.label = NULL,
        edge.curvature = pi/4,
        edge.col = ~ego(alpha=0, col = '#DB2B27') + alter(alpha = .7)),top='False Negative'),
  arrangeGrob(nplot(g_ad,
        layout = layout,
        vertex.label = NULL,
        edge.curvature = pi/4,
        edge.width = .8),left = 'Advice Network'),
  arrangeGrob(nplot(ad_1,
        layout = layout,
        vertex.label = NULL,
        edge.curvature = pi/4,
        edge.line.lty = 3,
        edge.width = .8)),
  arrangeGrob(nplot(ad_fp,
        layout = layout,
        vertex.label = NULL,
        edge.curvature = pi/4,
        edge.width = .8,
        edge.col = ~ego(alpha=0, col = '#55B748') + alter(alpha = .7))),
  arrangeGrob(nplot(ad_fn,
        layout = layout,
        edge.curvature = pi/4,
        edge.width = .8,
        edge.col = ~ego(alpha=0, col = '#DB2B27') + alter(alpha = .7))),
  ncol = 4, nrow = 2
)
# print(plot(gviz))
ggsave("netplot_v1.png",plot = gviz, width = 450, height = 250, units = 'mm') # take too long time to save it...
# dev.off()
```

## Calculate some basic stats for true graphs

```{r basic stats}
library(xtable)
ad.true <- as.network(apply(ad.post.actor$net, c(2, 3), median))
fr.true <- as.network(apply(fr.post.actor$net, c(2, 3), median))
# create a data.frame for a des stat table
table <- data.frame(Network = c('Actual Friendship Network','Perceived Friendship Networks', 'Actual Advice Network','Perceived Advice Networks'),
                    People = c(network.size(fr.true),paste(round(mean(sapply(friendship_nets,network.size)),2),'\u00B1',round(sd(sapply(friendship_nets,network.size)),2)),network.size(ad.true),paste(round(mean(sapply(advice_nets,network.size)),2),'\u00B1',round(sd(sapply(advice_nets,network.size)),2))),
                    Edges = c(network.edgecount(fr.true),paste(round(mean(sapply(friendship_nets,network.edgecount)),2),'\u00B1',round(sd(sapply(friendship_nets,network.edgecount)),2)),network.edgecount(ad.true),paste(round(mean(sapply(advice_nets,network.edgecount)),2),'\u00B1',round(sd(sapply(advice_nets,network.edgecount)),2))),
                    Density = c(network.density(fr.true),paste(round(mean(sapply(friendship_nets,network.density)),2),'\u00B1',round(sd(sapply(friendship_nets,network.density)),2)),network.density(ad.true),paste(round(mean(sapply(advice_nets,network.density)),2),'\u00B1',round(sd(sapply(advice_nets,network.density)),2))),
                    Mutual = c(dyad.census(fr.true)[[1,1]],paste(round(mean(sapply(friendship_nets, function(x) dyad.census(x)[[1,1]])),2),'\u00B1',round(sd(sapply(friendship_nets, function(x) dyad.census(x)[[1,1]])),2)),dyad.census(ad.true)[[1,1]],paste(round(mean(sapply(advice_nets, function(x) dyad.census(x)[[1,1]])),2),'\u00B1',round(sd(sapply(advice_nets, function(x) dyad.census(x)[[1,1]])),2))),
                    Asymmetric = c(dyad.census(fr.true)[[1,2]],paste(round(mean(sapply(friendship_nets, function(x) dyad.census(x)[[1,2]])),2),'\u00B1',round(sd(sapply(friendship_nets, function(x) dyad.census(x)[[1,2]])),2)),dyad.census(ad.true)[[1,2]],paste(round(mean(sapply(advice_nets, function(x) dyad.census(x)[[1,2]])),2),'\u00B1',round(sd(sapply(advice_nets, function(x) dyad.census(x)[[1,2]])),2))),
                    Null = c(dyad.census(fr.true)[[1,3]],paste(round(mean(sapply(friendship_nets, function(x) dyad.census(x)[[1,3]])),2),'\u00B1',round(sd(sapply(friendship_nets, function(x) dyad.census(x)[[1,3]])),2)),dyad.census(ad.true)[[1,3]],paste(round(mean(sapply(advice_nets, function(x) dyad.census(x)[[1,3]])),2),'\u00B1',round(sd(sapply(advice_nets, function(x) dyad.census(x)[[1,3]])),2))))

# reciprocity rate
grecip(ad.true) #.700
grecip(fr.true) #.919

xtable(table,include.rownames = FALSE)
```


## Rewiring

```{r rewiring}
library(netdiffuseR)

rewiring_css <- function(graph, nties = NULL){
  
  if (!length(nties))
    nties <- sapply(graph, sum)
  
  graph[2:length(graph)] <- lapply(2:length(graph), function(i) {
    as.matrix(rewire_graph(graph[[i]], p = 20 * nties[i], algorithm = "swap"))
    })
  
  graph <- blockdiagonalize(graph)
  
  graph <- new_barry_graph(as.matrix(graph), n = length(nties))

  out <- count_recip_errors(graph)
  out$name <- gsub("[(][0-9]+[)]", "", out$name)
  return(out)
}

#' Rewiring algorithm preserving degree-sequence
#' 
#' @param nsim Integer. Number of simulations
#' @param nets A list of CSS networks; the first one the original.
#' @param ncores Integer. Number of cores to run the test (parallel computing).
#' @param seed Integer. Seed for Pseudo RNG (passed to parallel::clusterSetRNGStream)
rewire_test <- function(
  nets,
  nsim   = 1000,
  ncores = 4,
  seed   = sample.int(.Machine$integer.max, 1)
  ) {

  # Preparing the data
  graph <- lapply(nets, as.matrix)
  graph[2:length(graph)] <- lapply(graph[2:length(graph)], methods::as, "dgCMatrix")
  nties <- sapply(graph, sum)

  # Setting up the cluster
  cl <- parallel::makePSOCKcluster(ncores)
  on.exit(parallel::stopCluster(cl)) # The cluster will stop automatically
  
  parallel::clusterSetRNGStream(cl, seed)
  
  parallel::clusterEvalQ(cl, {
    library(netdiffuseR)
    library(ergmito)
    library(imaginarycss)
  })
  
  parallel::clusterExport(cl, c("graph", "nties"), envir = environment())
  parallel::clusterExport(cl, "rewiring_css", envir = .GlobalEnv)
  
  # Making the call
  ans <- parallel::parLapply(cl, X = 1:nsim, fun = function(i) {
    rewiring_css(graph)
  })
  
  ans <- do.call(rbind, ans)
  return(ans)
  
}

# calculate empirical error rates
emp_count <- function(nets = nets){
  net0 <- new_barry_graph(as.matrix(blockdiagonalize(nets)),
                                  n = length(nets)-1)
  ans0 <- count_recip_errors(net0)
  ans0$name <- gsub("[(][0-9]+[)]", "", ans0$name)
  return(ans0)
}

# set BNAM as a true graph
# friendship 
nets <- lapply(advice_nets, as.matrix) # c(list(fr.true), friendship_nets)
nets <- c(list(nets[[1]]), nets)

fr.ans0 <- emp_count(nets = nets)
fr.ans <- rewire_test(nsim = 2000,
                        nets = nets)

# advice
nets <- c(list(ad.true), advice_nets)
ad.ans0 <- emp_count(nets = nets)

set.seed(1231) # For reproducibility
ad.ans2 <- rewire_test(nsim = 100,
                        nets = nets)
```

## Null distributions

```{r histograms}

ghist <- function(ans, ans0, id){
  library(ggplot2)
  hplot <- ggplot(data = ans[ans$id==id,],
       aes(x = value, group = name, fill = name)) +
    geom_histogram(alpha = .6)+
    geom_vline(data = ans0[ans0$id==id,],
               aes(xintercept = value, color = name),
               linetype = 2)+
      scale_fill_manual(values = c('#55B748','#DB2B27','#FDBF11','#98CF90','#E9807D')) +
      scale_color_manual(values = c('#55B748','#DB2B27','#FDBF11','#98CF90','#E9807D')) +
    facet_grid(name~.,
               labeller = labeller(name = as_labeller(c(`Partially false recip (omission) ` = 'P-FN',`Partially false recip (comission) ` = 'P-FP',`Completely false recip (omission) ` = 'C-FN',`Completely false recip (comission) ` = 'C-FP',`Mixed reciprocity errors ` = 'M-F'))))+
    theme_classic(base_size = 18) +
    theme(axis.title = element_text(face = "bold"),legend.position = 'none')+
    ggtitle(paste('Respondent',id+1,'False Motifs'))
  return(hplot)
}

# plot
for (id in 0:20){
  hplot <- ghist(ans = fr.ans,
                 ans0 = fr.ans0,
                 id = id)
  print(hplot)
}
```

```{r Bootstrap Hypothesis Testing}
bootstrap_testing <- function(meanobs,meansim){
  # n <- length(meanobs)
  ans <- vector("numeric",4)
  ans[3] <- mean(meanobs)
  ans[2] <- 2*min(mean(meansim <= ans[3]),
                  mean(meansim >ans[3]))
  ans[4] <- mean(meansim)
  ans[5] <- sd(meansim)
  # z-score
  ans[1] <- (ans[3]-ans[4])/ans[5]
  return(ans)
}

# id = c(0,20)
# class = c(1,5)
ind_test <- function(id = 0, class = 1, ans, ans0){
  ans$class <- as.numeric(as.factor(ans$name))
  ans0$class <- as.numeric(as.factor(ans0$name))
  # sapply(ans,levels)
  # sapply(ans0,levels)
  ans[,c('name','class')] # check which name corresponds to class numeric
  # define meansim as a simulated values vector
  meansim <- ans[ans$id==id & ans$class==class,'value']
  # define meanobs as a observed value
  meanobs <- ans0[ans0$id==id & ans0$class==class,'value']
  
  # now run the bootstrap hypothesis test
  ans <- bootstrap_testing(meanobs = meanobs, meansim = meansim)
  return(ans)
}

# aggregated class test
agg_test <- function(class = 1, ans, ans0){
  ans$class <- as.numeric(as.factor(ans$name))
  ans0$class <- as.numeric(as.factor(ans0$name))
  meanobs <- ans0[ans0$class==class,'value']
  meansim <- ans[ans$class==class,'value']
  
  # now run the bootstrap hypothesis test
  out <- bootstrap_testing(meanobs = meanobs, meansim = meansim)
  return(out)
}
ind_test(id = 20, class = 5, ans = fr.ans, ans0 = fr.ans0)
t(sapply(c(0:20),ind_test,1,fr.ans,fr.ans0))
fr.table <- cbind(t(sapply(c(0:20),ind_test,1,fr.ans,fr.ans0)),
      t(sapply(c(0:20),ind_test,2,fr.ans,fr.ans0)),
      t(sapply(c(0:20),ind_test,3,fr.ans,fr.ans0)),
      t(sapply(c(0:20),ind_test,4,fr.ans,fr.ans0)),
      t(sapply(c(0:20),ind_test,5,fr.ans,fr.ans0)))
colnames(fr.table) <- rep(c('Z' , 'p','obs','M','SD'),5)
# the following function comes from https://stackoverflow.com/questions/13984470/possible-to-create-latex-multicolumns-in-xtable
construct_header <- function(df, grp_names, span, align = "c", draw_line = T) {
  if (length(align) == 1) align <- rep(align, length(grp_names))
  if (!all.equal(length(grp_names), length(span), length(align)))
    stop("grp_names and span have to have the same length!")

  if (ncol(df) < sum(span)) stop("Span has to be less or equal to the number of columns of df") 

  header <- mapply(function(s, a, grp) sprintf("\\multicolumn{%i}{%s}{%s}", s, a, grp),
                   span, align, grp_names)
  header <- paste(header, collapse = " & ")
  header <- paste0(header, " \\\\")

  if (draw_line) {
    # where do we span the lines:
    min_vals <- c(1, 1 + cumsum(span)[1:(length(span) - 1)])
    max_vals <- cumsum(span)
    line <- ifelse(grp_names == "", "", 
                   sprintf("\\cmidrule(lr){%i-%i}", min_vals, max_vals))
    line <- paste(line[line != ""], collapse = " ")

    header <- paste0(header, "  ", line, "\n  ")
  }

  addtorow <- list(pos = list(-1, -1, nrow(df)),
                   command = c("\\hline\n  ", header, "\\hline\n  "))
  return(addtorow)
}

a_header <- construct_header(
  fr.table,
  grp_names = c('','C-FP','C-FN','M-F','P-FP','P-FN'),
  span = c(5,5,5,5,5),
  align = 'c'
)
print(xtable(fr.table), add.to.row = a_header, scalebox = .5)
# advice table
ad.table <- cbind(t(sapply(c(0:20),ind_test,1,ad.ans,ad.ans0)),
      t(sapply(c(0:20),ind_test,2,ad.ans,ad.ans0)),
      t(sapply(c(0:20),ind_test,3,ad.ans,ad.ans0)),
      t(sapply(c(0:20),ind_test,4,ad.ans,ad.ans0)),
      t(sapply(c(0:20),ind_test,5,ad.ans,ad.ans0)))
colnames(ad.table) <- rep(c('Z','p','obs','M','SD'),5)
a_header <- construct_header(
  ad.table,
  grp_names = c('C-FP','C-FN','M-F','P-FP','P-FN'),
  span = c(5,5,5,5,5),
  align = 'c'
)
print(xtable(ad.table), add.to.row = a_header, scalebox = .52)

# aggregate test
agg.table <- cbind(t(sapply(1:5,agg_test,ans = fr.ans, ans0 = fr.ans0)),
                   t(sapply(1:5,agg_test,ans = ad.ans, ans0 = ad.ans0)))
agg_header <- construct_header(
  agg.table,
  grp_names = c('Friendship Network','Advice Network'),
  span = c(5,5),
  align = 'c'
)
print(xtable(agg.table), add.to.row = agg_header, scalebox = .57)
```

Using ans, we visualize histograms to check the distributions of each error case. Here, we use ggplot2.

```{r visualization}
library(ggplot2)
library(dplyr)

g_hist2 <- function(ans, ans0){
  ghist2 <- ggplot(data = ans0,
        aes(x = value, fill=name))+
  geom_histogram(alpha=.6)+
  geom_vline(data=ans0 %>% group_by(name) %>% mutate(avg=mean(value)), aes(xintercept = avg, color=name),
             linetype = 2)+
  scale_fill_manual(values = c('#55B748','#DB2B27','#FDBF11','#98CF90','#E9807D')) +
  scale_color_manual(values = c('#55B748','#DB2B27','#FDBF11','#98CF90','#E9807D')) +
  geom_text(data=ans0 %>% group_by(name) %>% mutate(avg=round(mean(value),2)), mapping = aes(x = Inf, y = Inf, label = paste0('M = ',avg), color = name), hjust = 1.05, vjust = 1.5)+
  facet_grid(name~.,
             labeller = labeller(name = as_labeller(c(`Partially false recip (omission) ` = 'P-FN',`Partially false recip (comission) ` = 'P-FP',`Completely false recip (omission) ` = 'C-FN',`Completely false recip (comission) ` = 'C-FP',`Mixed reciprocity errors ` = 'M-F'))))+
  theme_classic(base_size = 14) +
  theme(axis.title = element_text(face = "bold"),legend.position = 'none')
  return(ghist2)
}

hist_save <- grid.arrange(arrangeGrob(g_hist2(ans = fr.ans,ans0 = fr.ans0), top = '(a) Friendship Network'),
             arrangeGrob(g_hist2(ans = ad.ans,ans0 = ad.ans0), top = '(b) Advice Network'),
             ncol = 2, nrow = 1)
ggsave('hist_agg_v1.png', plot = hist_save, dpi = 800)
```

## Correration Analysis

Here, we analyze different class of false patterns based on individual attributes.

```{r regression}
library(ggpubr)
attr_df <- read.csv('../data/HIghTech_attr.csv',sep = ';')
attr_df$fr_indegree <- degree(fr.true, cmode = 'indegree')
attr_df$fr_outdegree <- degree(fr.true, cmode = 'outdegree')
attr_df$ad_indegree <- degree(ad.true, cmode = 'indegree')
attr_df$ad_outdegree <- degree(ad.true, cmode = 'outdegree')
attr_df$fr_density <- sapply(friendship_nets,network.density)
attr_df$ad_density <- sapply(advice_nets,network.density)

fr_df <- data.frame(rbind(t(sapply(c(0:20),ind_test,1,fr.ans,fr.ans0)),
      t(sapply(c(0:20),ind_test,2,fr.ans,fr.ans0)),
      t(sapply(c(0:20),ind_test,3,fr.ans,fr.ans0)),
      t(sapply(c(0:20),ind_test,4,fr.ans,fr.ans0)),
      t(sapply(c(0:20),ind_test,5,fr.ans,fr.ans0))))
colnames(fr_df) <- c('Z','p','obs','sim.M','sim.SD')
fr_df$id <- rep(1:21,5)
fr_df$class <- c(rep('C-C',21),rep('C-O',21),rep('M-F',21),rep('P-C',21),rep('P-O',21)) #this is only for the plot order
fr_df$Z[is.nan(fr_df$Z)] <- 0

ad_df <- data.frame(rbind(t(sapply(c(0:20),ind_test,1,ad.ans,ad.ans0)),
      t(sapply(c(0:20),ind_test,2,ad.ans,ad.ans0)),
      t(sapply(c(0:20),ind_test,3,ad.ans,ad.ans0)),
      t(sapply(c(0:20),ind_test,4,ad.ans,ad.ans0)),
      t(sapply(c(0:20),ind_test,5,ad.ans,ad.ans0))))
colnames(ad_df) <- c('Z','p','obs','sim.M','sim.SD')
ad_df$id <- rep(1:21,5)
ad_df$class <- c(rep('C-C',21),rep('C-O',21),rep('M-F',21),rep('P-C',21),rep('P-O',21))
ad_df$Z[is.nan(ad_df$Z)] <- 0

twovars_plot <- function(df = fr_df, 
                         var1 = fr_outdegree, 
                         var2 = Z, 
                         var1_name = 'Outdegree', 
                         var2_name = 'Z-score'){
  splot <- left_join(df,
          attr_df,
          by = c('id' = 'id')) %>% 
  ggplot(data = .,
         aes(x = !!enquo(var1),y = !!enquo(var2), color = class))+
  geom_point(alpha = .6)+
  geom_smooth(method = 'lm')+
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01,hjust = 0, vjust = 1)+
  scale_color_manual(values = c('#55B748','#DB2B27','#FDBF11','#98CF90','#E9807D')) +
  facet_grid(class~.,
             labeller = labeller(class = as_labeller(c(`C-C`='C-FP',`C-O`='C-FN',
                                                      `M-F`='M-F',
                                                      `P-C`='P-FP',
                                                      `P-O`='P-FN')))) +
  theme_classic(base_size = 14) +
  theme(axis.title = element_text(face = "bold"),legend.position = 'none')+
  ylab(var2_name)+
  xlab(var1_name)
  return(splot)
}

outdegree <- grid.arrange(arrangeGrob(twovars_plot(),top = '(a) Friendship Network'),
             arrangeGrob(twovars_plot(df = ad_df,var1=ad_outdegree),top = '(b) Advice Network'),
             ncol = 2, nrow = 1)
ggsave('outdegree_cor_v1.png', plot = outdegree, dpi = 800)

den_plot <- grid.arrange(arrangeGrob(twovars_plot(var1 = fr_density,var1_name = 'Density'),top = '(a) Friendship Network'),
             arrangeGrob(twovars_plot(df=ad_df,var1 = ad_density,var1_name = 'Density'),top = '(b) Advice Network'),
             ncol = 2, nrow = 1)
ggsave('density_cor_v1.png', plot = den_plot, dpi = 800)

lm(Z~fr_density+tenure+factor(hierarchy)+factor(dept),
   left_join(fr_df,
          attr_df,
          by = c('id' = 'id')) %>% .[.$class=='C-C',]) %>% summary
```

